We present a system designed for visualizing and analyzing multivariate time series. This system is built on a novel framework based on multiscale continuous time Markov chains which is extracted from given data. The time-series can then be visualized via a graph along with auxiliary visualziations to help us understand the temporal and spatial aspects of each state and transition. The system provides automatic naming and rule extraction services in addition to the individual visualizations.  

We present several examples of analysis conducted with the system.  There are several directions for further work. 
%\begin{itemize}

\noindent  {\bf Online Modeling:} Currently the multiscale model is fixed based on initial data. Although new data can be mapped onto the model to provide services such as alerts and basic anomoly detection in real time, an open question is how to learn and update the model for streaming data efficiently. This would require efficient schemes for updating the multiscale model and the transition probabilities. This would also provide opprotunities for visualizing changes in the structure of the Markov chains.

\noindent {\bf Automatic scale selection:} StreamStory provides an interactive visualization of the multiscale structure, however not all the scales are equally informative.  We have begun initial work on selecting a several discrete scales to present to the user automatically. These are based on functional description of the hierarchical structure of the Markov chains. This feature has not been fully integrated into the system but would provide an additional tool to help distill the  information presented to the user. 

\noindent {\bf Reconstruction:} We have not discussed reconstruction results for our system. One direction would be to establish conditions under which we can reconstruct the ``correct'' dynamics of the underlying system. The connection with dynamical systems theory remains largely unexplored.

\noindent {\bf Aggregation Schemes:} On our tested datasets, the two aggregations schemes for computing the multiscale hierarchy produced similar results. We plan to investigate if there is a deeper connection between the approaches and undderstand what their relationship is. 

%\end{itemize}
%discussion: reconstruction results
%  %                   continuous state space


% differences in aggreagtion schemes



% \begin{itemize}
% 	\item Kaj se zgodi, ce model updates v real time?
% 	\item Automatic description generation.
% 	\item theoretical guarantees
% 	\item  
% \end{itemize}